### 2021-04-15 submission/manuscript.Rmd
Add mice per timepoint range to Fig 4 legend and update Fig 4 in manuscript

library(tidyverse)
library(knitr)
library(here)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=2, nsmall=2, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)
```

@@

``` {r figure_4_permnova}
one_day_permnova_results <- read_tsv(here("data/process/1_day_PEG_permanova.tsv"))
one_day_top_2_contrib <- one_day_permnova_results %>% arrange(desc(R2)) %>%
  top_n(3, R2)
one_day_top_contrib_name = (one_day_top_2_contrib[2,1])
one_day_sec_contrib_name = (one_day_top_2_contrib[3,1])
top_R2 = one_day_top_2_contrib[2,6]
sec_R2 = one_day_top_2_contrib[3,6]
```
\includegraphics{figure_4.pdf}
**Figure 4. 1-day PEG treatment renders mice susceptible to transient *C. difficile* colonization.** A. Setup of the experimental timeline for the 1-day PEG treated subset of mice. B. CFU/gram stool measured over time (N = 6 mice per timepoint) via several dilutions. The black dotted line represents the limit of detection for the first serial dilution. Asterisks indicate timepoints where the CFU/gram was significantly different between groups using the Kruskall-Wallis test with a Benjamini-Hochberg correction for multiple timepoints. C. Principle Coordinate Analysis plot of the groups over time with the alpha representating the same time scale as in panel D (`r one_day_top_contrib_name`: R^2^ = `r top_R2`; `r one_day_sec_contrib_name`: R^2^ = `r sec_R2`). D. Shannon diversity Index of the groups over time. Only days with samples from all groups are shown. Samples for some mice were difficult to obtain 
due to the laxative treatment. The alpha scale follows accordingly with the timeline. E. Line plots of relative percent abundance of selected genera over time. Only days with samples from all groups shown. The gray line represents the limit of detection.

>>>>

library(tidyverse)
library(knitr)
library(here)
library(readxl)

@@

	if(is.numeric(x)){
		if(abs(x - round(x)) < .Machine$double.eps^0.5){
			paste(format(x,big.mark=',', digits=0, scientific=FALSE))
		} else {
			paste(format(x,big.mark=',', digits=2, nsmall=2, scientific=FALSE))
		}
	} else {
    	paste(x)      
	}
}
knitr::knit_hooks$set(inline=inline_hook)

metadata <- read_excel(here("data/process/metadata.xlsx"), col_types = c("text", "numeric", "text", "text", "numeric", "text", "numeric", "text", "text", "text", "text", "numeric", "numeric")) #specify column types

@@

``` {r figure_4_permnova}
one_day_permnova_results <- read_tsv(here("data/process/1_day_PEG_permanova.tsv"))
one_day_top_2_contrib <- one_day_permnova_results %>% arrange(desc(R2)) %>%
  top_n(3, R2)
one_day_top_contrib_name = (one_day_top_2_contrib[2,1])
one_day_sec_contrib_name = (one_day_top_2_contrib[3,1])
top_R2 = one_day_top_2_contrib[2,6]
sec_R2 = one_day_top_2_contrib[3,6]

one_day_samples_per_day <- metadata %>%
  filter(group == "C" & exp_num %in% c("M6")| #Only use C mice from this experiments. Allocated groups to figures based on paper outline.
             group == "1RM1" & exp_num %in% c("M6R")| #Had to differentiate experiment 6 from 6R in the metadata to create unique_mouse_id that wouldn't overlap for the M1 & 1RM1 mice that are both labeled with mouse_ids that are #s1-6
             group == "M1" & exp_num %in% c("M6")) %>%
  group_by(day) %>%
  count() 
one_day_sample_min <- min(one_day_samples_per_day$n)
one_day_sample_max <- max(one_day_samples_per_day$n)

```
\includegraphics{figure_4.pdf}
**Figure 4. 1-day PEG treatment renders mice susceptible to transient *C. difficile* colonization.** A. Setup of the experimental timeline for the 1-day PEG treated subset of mice. B. CFU/gram stool measured over time (N = `r one_day_sample_min`-`r one_day_sample_max` mice per timepoint) via several dilutions. The black dotted line represents the limit of detection for the first serial dilution. Asterisks indicate timepoints where the CFU/gram was significantly different between groups using the Kruskall-Wallis test with a Benjamini-Hochberg correction for multiple timepoints. C. Principle Coordinate Analysis plot of the groups over time with the alpha representating the same time scale as in panel D (`r one_day_top_contrib_name`: R^2^ = `r top_R2`; `r one_day_sec_contrib_name`: R^2^ = `r sec_R2`). D. Shannon diversity Index of the groups over time.  Only days with samples from all groups are shown. E. Line plots of relative percent abundance of selected genera over time. Only days with samples from all groups shown. The gray line represents the limit of detection.

### 2021-04-15 submission/manuscript.tex
Add mice per timepoint range to Fig 4 legend and update Fig 4 in manuscript

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since laxatives and
clindamycin have previously been shown to disrupt the murine microbiota
(ref), we hypothesized the different \emph{C. difficile} colonization
dynamics between mice treated with laxatives versus clindamycin were due
to their differential effects on the microbiota. We profiled the stool
microbiota over time by sequencing the V4 region of the 16S rRNA gene to
compare changes across treatment groups. We found time and treatment
group explained half of the observed variation between fecal communities
(PERMANOVA combined R\textsuperscript{2} = 0.50, \emph{P} \textless{}
0.001).

@@

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental timeline for the 1-day PEG treated subset
of mice. B. CFU/gram stool measured over time (N = 6 mice per timepoint)
via several dilutions. The black dotted line represents the limit of
detection for the first serial dilution. Asterisks indicate timepoints
where the CFU/gram was siginificantly different between groups using the
Kruskall-Wallis test with a Benjamini-Hochberg correction for multiple
timepoints. C. Principle Coordinate Analysis plot of the groups over
time with the alpha representating the same time scale as in panel D
(day: R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} = 0.19).
D. Shannon Diveristy Index of the groups over time. Only days with
samples from all groups are shown. Samples for some mice were difficult
to obtain due to the laxative treatment. The alpha scale follows
accordingly with the timeline. E. Line plots of relative percent
abundance of selected genera over time. Only days with samples from all
groups shown. The gray line represents the limit of detection.

>>>>

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since laxatives and
clindamycin have previously been shown to disrupt the murine microbiota
(ref), we hypothesized the different \emph{C. difficile} colonization
dynamics between mice treated with laxatives versus clindamycin were due
to their differential effects on the microbiota. We profiled the stool
microbiota over time by sequencing the V4 region of the 16S rRNA gene to
compare changes across treatment groups. We found time and treatment
group explained half of the observed variation between fecal communities
(PERMANOVA combined R\textsuperscript{2} = 0.50, \emph{P} \textless{}
0.001) with the remaining variation explained by interactions between
treatment group and other experimental variables including time, cage
effects, and sequencing preparation (Data Set S1, Sheet X). The impact
of cage effects has been well documented due to copraphagy and is
expected.

@@

\includegraphics{figure_4.pdf} \textbf{Figure 4. 1-day PEG treatment
renders mice susceptible to transient \emph{C. difficile} colonization.}
A. Setup of the experimental timeline for the 1-day PEG treated subset
of mice. B. CFU/gram stool measured over time (N = 12-18 mice per
timepoint) via several dilutions. The black dotted line represents the
limit of detection for the first serial dilution. Asterisks indicate
timepoints where the CFU/gram was significantly different between groups
using the Kruskall-Wallis test with a Benjamini-Hochberg correction for
multiple timepoints. C. Principle Coordinate Analysis plot of the groups
over time with the alpha representating the same time scale as in panel
D (day: R\textsuperscript{2} = 0.43; group: R\textsuperscript{2} =
0.19). D. Shannon diversity Index of the groups over time. Only days
with samples from all groups are shown. E. Line plots of relative
percent abundance of selected genera over time. Only days with samples
from all groups shown. The gray line represents the limit of detection.
